---
title: "Setting Up Zenodo"
author: "Daniel Antal, CFA"
date: "5/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In our observatories, we store the authentic copies of new datasets on the European scientific data repository, [Zenodo](https://zenodo.org/). If you are new to Zenodo, you should upload at least one or two datasets manually, before trying to automate the process. And to avoid live-testing in Zenodo, where everything is supposed to be permanent, set up a practice account on its practice clone,  [Sandbox Zenodo](https://sandbox.zenodo.org/) account.  

In this example, you will authenticate yourself twice:  you will authenticate yourself as the creator (author) of a scientific object with [ORCID](https://orcid.org/). ORCID provides a persistent digital identifier (an ORCID ID) that you own and control, and that distinguishes you from other researchers. And you will authenticate your session to [Zenodo](https://zenodo.org/) /  [Sandbox Zenodo](https://sandbox.zenodo.org/)  with a personal access token to one of these sites.  Note that your Zenodo credentials will not work on Sandbox Zenodo, which ensures that you cannot accidentally upload practicing material to the real Zenodo, where there is no `undo` button.

## Set Up Your Zenodo / Sandbox Account

For practicing, please set up a [Zenodo Sandbox](https://sandbox.zenodo.org/) account. Zenodo Sandbox is a clone of Zenodo, created for testing applications. Do not practice on Zenodo itself, because whatever you publish on Zenodo is permanent. All examples below work exactly the same way on Zenodo, without the `sandbox` subdomain in the calls.

Important: you will get a verification message on your account and your email address.  If you do not reply to the verification links (check Spam, Social, etc.) the API will seemingly work, but it will not record anything. That will lead to misleading error messages.

Once you are sure that your Sandbox account is up and running (and verified), you should create your Personal Access Token (PAT). In your user profile, go to `Applications` and create a secret code by clicking deposit:actions, deposit:write.  

In R, the best practice is to store this PAT in a keyring with the [keyring](https://r-lib.github.io/keyring/index.html) package. The following code will interactively set your PAT, i.e. if you run it in R, a pop-up window will ask you to copy the Zenodo Sandbox PAT from your browser to a textbox. Of course, you can use your favorite method [managing your secret variables](https://cran.r-project.org/web/packages/httr/vignettes/secrets.html), but do not accidentally upload your PAT to github or send it to somebody in an email. If you store it in your repo, make sure that you add it to `.gitignore` file, and in a package add it to `.Rbuildignore`, too.

```{r, eval=FALSE}
require(keyring)
keyring::key_set (service =  "Zenodo_Sandbox", username = NULL)  # separately for the sandbox
keyring::key_set (service =  "Zenodo", username = NULL)          # and the real service
```

## Your First Deposition

Your deposition has three important parts:

* Metadata set for unique identification by Zenodo (Sandbox), such as a Zenodo ID and a digital object identifier (DOI), or a version of a DOI.
* Metadata set by the user, including the metadata of your authorship, type of the document, keywords, etc.  Use ORCID to authenticate yourself. 
* A copy of the data in a file. 

In this tutorial we use [Zen4R](https://github.com/eblondel/zen4R/wiki), the R Interface to the Zenodo REST API. Zen4R uses R6 objects to prepare your deposition, which, unless you are familiar with object-oriented languages, may seem a bit bizarre at first. R6 objects are not real R objects, but environments, so when you create a metadata record they will not show in RStudios's Environment window as an object, but as a new environment. This should not be a problem, but if you experience unexpected behavior, be mindful that you are not debugging a true R object in your global or function environment. 

### Start New Session 

Using the keyring you you set up earlier, you initiate a new session with the API:

```{r new-zenodo-session}
require(zen4R)    # for Zenodo API interaction 
require(keyring)  # you can use any other secure form to store your PAT

ZENODO <- ZenodoManager$new(
url = "https://sandbox.zenodo.org/api",
token = keyring::key_get( "Zenodo_Sandbox"),
logger = "INFO"
)

```

If you did not store your PAT anywhere, you can also simply write `token = 'abc_mytoken_def'` where 'abc_mytoken_def' is of course your secret token generate on [Zenodo Sandbox](https://sandbox.zenodo.org/). If you have failed to save, you can always generate a new one in the web interface.

### Create a Record

In this step we are creating an R6 object called `myrec`. If you run this code, in your RStudio, in the Environment window you will see `myrec` not as an object, but as an `Environment`.

```{r creat-fill-record, echo=FALSE}
#create and fill a local (not yet deposited) record
myrec <- ZenodoRecord$new()
myrec$setTitle("Internet use: participating in social networks [percentage of individuals]")
myrec$setDescription("Digital Music Observatory, re-processed Eurostat data")
myrec$setUploadType("dataset")
myrec$addCreator(
  firstname = "Jane", lastname = "Doe",
  affiliation = "Independent", orcid = "0000-0001-0000-0000"
 )
myrec$setLicense("mit")
myrec$setAccessRight("open")
myrec$setDOI("http://doi.org/00.0000/zenodo.00000") #use this method if your DOI has been assigned elsewhere, outside Zenodo

```

If you run `myrec` in your console, you will be able to print out your record, but only if you replaced `Jane Doe` with your own name, and `orcid` with your true `ORCID ID`.  Also, if you specify a pre-set DOI (and don't expect to get a new DOI from Zenodo), the API will check the validity of the `DOI` and the `ORID ID`, `firstname`, `lastname`. So you cannot test the following code with the dummy `http://doi.org/00.0000/zenodo.00000` and `Jane Doe`.  

Because we work with `R6` objects, your session is called `ZENODO`, and you call the `depositRecord` method on session `ZENODO` to assign the `myrec` ZenodoRecord record object. With `Jane Doe`, you will get a validation error, but it should work fine, provided that you have spelled your name the way it stands in your [ORCID](https://orcid.org/) ID records. 

```{r get-error}
myrec <- ZENODO$depositRecord(myrec)
```
If you have correctly provided real information, have gone through the verification process, take a look at your [Zenodo Sandbox](https://sandbox.zenodo.org/) account's `uploads`,  you should see a record, but not the data.

### Upload the Data

Crucially, at this point your record should have a Zenodo ID that connects your ORCID ID with the metadata record. In this example, because we used `Jane Doe`, you get an NULL id back. 

```{r recorid-d}
myrec$id
```

Should you work programmatically, your script can check for empty IDs before the upload:

```{r just-a-test}
is.null(myrec$id)
```

A real-life example must return `FALSE`. The final step is to connect your `ZENODO` session (the `R6` object) to an R file that should be uploaded. You could upload `rds` files to Zenodo, but I would suggest something more system- and language independent, such as `csv` or `json`.  

In this example, I create a temporary file in your R session, I write there the famous `iris` dataset in csv format, and try to upload it to the `ZENODO` session.  In this case, you get an error  because `Jane Doe` with her fictitious ORCID ID and non-existing DOI was prevented from spamming the server.

```{r upload-the-file}
my_file_path <- tempfile()
write.csv(iris, my_file_path, row.names = FALSE)
ZENODO$uploadFile(my_file_path, myrec$id)
```


## Your Authoritative Copy

You can check what I uploaded here on [sandbox.zenodo.org/deposit/818354](https://sandbox.zenodo.org/deposit/818354). This is a copy of one of my datasets that you can find with full description on Zenodo, under the same DOI.
